---
title: 'Unnesting project: Starwars'
author: "Jonathan Mulamba"
date: "2024-06-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Why unnest data?

Unnesting is a crucial step in data preparation , ensuring that the data is in a usable format for analysis and visualization.
In this project , I unnest the starwars dataset from the tidyverse library and thereafter perform data analysis and create visualization.


**Loading the tidyverse & dplyr package and setting the theme for the visualizations to follow**
```{r library}
library(tidyverse)
library(dplyr)
theme_set(theme_grey())
```

**Utilizing the glimpse function provided by the 'dplyr' package to get a quick overview of the structure and content of the dataframe.**
```{r glimpse}
glimpse(starwars)
```
The glimpse function shows in the starwars dataset that the following columns are list datatypes/data structure: films, vehicles & starships.

Creating a new data frame to focus on the columns that we require for unnesting purposes.
```{r}
starwars_sm <- starwars %>% 
  select(name, 
         films,
         vehicles,
         starships)
```

#### Unnesting the films column
```{r}
sw_films <- starwars_sm %>% 
  unnest(films)

view(sw_films)
head(sw_films)
```

### Data analysis
Count the number of times each character appeared in a movie.
```{r , echo= FALSE}
sw_films %>% 
  count(name) %>% 
  arrange(-n)
```

### Data visulization
Create a bar chart to show the characters and the number of movies they appeared in
```{r pressure, echo=FALSE}
sw_films %>% 
  count(name) %>% 
  filter(n >= 4) %>% 
  mutate(name = fct_reorder(name, n)) %>% 
  ggplot(aes(y = name,
             x = n )) +
  geom_col()
```



#### Unnesting the starship column
```{r}
sw_ships <- starwars_sm %>% 
  unnest(starships, keep_empty =  TRUE)
head(sw_ships)
```


### Unnesting multiple columns
```{r}
sw_long <- starwars_sm %>% 
  unnest(starships) %>% 
  unnest(vehicles)

head(sw_long)
```

